---
title: "Review"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan <i>adapted from slides by Hastie & Tibshirani</i>
</span>
</div> 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ISLR)
library(tidymodels)
library(gridExtra)
```

---

class: center, middle

## RStudio Cloud Update


---

class: center, middle

## Make sure to bring a calculator to the exam on Thursday

---


## Office Hours this week:

* Wednesday 10a
* **Wednesday 2p**
* Thursday 10a

---

mpg | weight | automatic 
----|--------|-----
21  | 2.6 | 1
22.8 | 2.32 | 1
18.7 | 3.4 | 0
14.3 | 3.5 | 0
24.4 | 3.1 | 0

You want to predict mpg using weight and automatic. Write out how you would calculate $\hat\beta$ in matrix form using the data provided (you need not **solve** the matrices)

---

mpg | weight | automatic 
-----|--------|-----
21  | 2.6 | 1
22.8 | 2.32 | 1
18.7 | 3.4 | 0
14.3 | 3.5 | 0
24.4 | 3.1 | 0

Solving the equation results in the following $\hat\beta$:

$\begin{bmatrix} \hat{\beta}_0 \\ \hat{\beta}_1 \\ \hat{\beta}_2\end{bmatrix} = \begin{bmatrix}80.45 \\ -18.40 \\ -13.30\end{bmatrix}$


Using the information provided, calculate the MSE (mean squared error) for this model.

---

mpg | weight | automatic 
-----|--------|-----
21  | 2.6 | 1
22.8 | 2.32 | 1
18.7 | 3.4 | 0
14.3 | 3.5 | 0
24.4 | 3.1 | 0

$\begin{bmatrix} 0.66 &  0.34& 0.07& 0.19&-0.26 \\ 0.34& 0.66&-0.07&-0.19& 0.26 \\ 0.07&-0.07& 0.37& 0.42& 0.21 \\ 0.19&-0.19& 0.42& 0.55& 0.02 \\ -0.26& 0.26& 0.21& 0.02& 0.77 \\ \end{bmatrix}$

Above is the **hat matrix** for the model. Use this to calculate LOOCV error for this model

---

mpg | weight | automatic 
----|--------|-----
8  | 3.8 | 0
30.3 | 1.9 | 1
40.9 | 1.5 | 1
22.1 | 3.4 | 0
26 | 2.1 | 1

You get a new test data set (above). Using the same model you fit to the training data, calculate the MSE in this test dataset

---

class | x
-------|---
1 | -0.57
1 | -1.26
1 | 1.13
2 | 3.3
2 | 5.4

Using the data above, calculate $\pi_k$, $\mu_k$, and $\sigma^2$

---

class | x
-------|---
1 | -0.57
1 | -1.26
1 | 1.13
2 | 3.3
2 | 5.4

Using LDA, what is are the discrimant scores for x = 3? Which class would you classify this point into?

---

class | x
-------|---
1 | -0.57
1 | -1.26
1 | 1.13
2 | 3.3
2 | 5.4

What is the probability that an observation where x = 3 is in class 1?

---

class | x
-------|---
1 | -0.57
1 | -1.26
1 | 1.13
2 | 3.3
2 | 5.4

What is the decision boundary for this LDA problem?

---

## P(+ | Drug User) = 0.8
## P(Drug User) = 0.1
## P(+ | Not a drug user) = 0.2

* What is the probability of being a drug user give you test positive?
---

| Truth (+) | Truth (-) | total
-|---------|-------|
Predicted (+) | 10 | 20 | 30
Predicted (-) | 5 | 50 | 55
Total | 15 | 70 | 85

* What is the accuracy?
* What is the false positive rate?
* What is the false negative rate?