---
title: "CV Lab"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan <i>adapted from slides by Hastie & Tibshirani</i>
</span>
</div> 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(ISLR)
```

---

## RStudio Pro

# rstudio.hpc.ar53.wfu.edu:8787

## pw: R2D2Star!

---

class: center, middle

## Practice getting your lab into RStudio Pro

---

class: center, middle

## Now get your lab into RStudio Cloud

---

## Practice pushing and pulling in GitHub

* So far we have just been **pushing** to GitHub, but a nice feature is that you can then **pull** your results at any time. 
* Make a change to your Lab file in RStudio Pro and **push** this to GitHub (check the GitHub repo to make sure your changes showed up)
* Check RStudio cloud - did those changes show up?

--
* Click the pull (down) arrow in RStudio cloud. Do you see the changes now?

---

## Some R functions

```{r}
sample(10)
```

```{r}
sample(10)
```

---


## Some R functions

```{r}
set.seed(1)
sample(10)
```

```{r}
set.seed(1)
sample(10)
```

---

class: center, middle

## Use the `set.seed()` function before performing cross validation so you get consistent (reproducible) answers!

---

## Polynomial regression in R

```{r, echo = FALSE}
data <- data.frame(x = rnorm(100), y = rnorm(100))
```

$y = \beta_0 + \beta_1 x + \beta_2 x^2 + \beta_3 x^3 + \epsilon$

.small[
```{r}
lm(y ~ x + I(x^2) + I(x^3), data = data)
```
]

--

.small[
```{r}
lm(y ~ poly(x, 3), data = data)
```
]

---

## Polynomial regression in R

$y = \beta_0 + \beta_1 x + \beta_2 x^2 + \beta_3 x^3 + \epsilon$

```{r}
model1 <- lm(y ~ x + I(x^2) + I(x^3), data = data)
model2 <- lm(y ~ poly(x, 3), data = data)

all.equal(predict(model1), predict(model2))
```

---

class: center, middle

## When doing polynomial regression, you can use the `poly()` function in R

