---
title: "Lasso and Elastic Net"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan <i>adapted from slides by Hastie & Tibshirani</i>
</span>
</div> 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(ISLR)
library(countdown)
```

---

## Ridge Review

.question[
What are we minimizing with Ridge Regression?
]

--

$$RSS + \lambda\sum_{j=1}^p\beta_j^2$$

--

.question[
What is the resulting estimate for $\hat\beta_{ridge}$?
]

--

$$\hat\beta_{ridge} = (\mathbf{X}^{T}\mathbf{X}+\lambda\mathbf{I})^{-1}\mathbf{X}^T\mathbf{y}$$

--

.question[

Why is this useful?
]

---

## Ridge Review

.question[
How is $\lambda$ determined?
]

$$RSS + \lambda\sum_{j=1}^p\beta_j^2$$

--

.question[
What is the bias-variance trade-off?
]

---

## Ridge Regression

.pull-left[

## Pros

* Can be used when $p > n$
* Can be used to help with multicollinearity
* Will decrease variance
(as $\lambda \rightarrow \infty$ )

]

--

.pull-right[

## Cons

* Will have increased bias (compared to least squares)
* Does not really help with variable selection (all variables are included in _some_ regard, even if their $\beta$ coefficients are really small)

]

---

## Lasso!

* The lasso is similar to ridge, but it actually drives some $\beta$ coefficients to 0! (So it helps with variable selection)

--

$$RSS + \lambda\sum_{j=1}^p|\beta_j|$$

--

* We say lasso uses an $\ell_1$ penalty, ridge uses an $\ell_2$ penalty
--

* $||\beta||_1=\sum|\beta_j|$
* $||\beta||_2=\sum\beta_j^2$

---

## Lasso

* Like Ridge regression, lasso shrinks the coefficients towards 0
--

* In lasso, the $\ell_1$ penalty **forces** some of the coefficient estimates to be **exactly zero** when the tuning parameter $\lambda$ is sufficiently large
--

* Therefore, lasso can be used for **variable selection**
--

* The lasso can help create **smaller, simplier** models
--

* Choosing $\lambda$ again is done via cross-validation

---

## Ridge Regression

.pull-left[

## Pros

* Can be used when $p > n$
* Can be used to help with multicollinearity
* Will decrease variance
(as $\lambda \rightarrow \infty$ )
* Can be used for variable selection, since it will make some $\beta$ coefficients exactly 0

]

--

.pull-right[

## Cons

* Will have increased bias (compared to least squares)
* If $p>n$ the lasso can select **at most** $n$ variables

]

---

## Ridge versus lasso

* Neither Ridge nor lasso will universally dominate
--

* Cross-validation can also be used to determine which method (Ridge or lasso) should be used
--

* Cross-validation is **also** used to select $\lambda$ in either method. You choose the $\lambda$ value for which the cross-validation model is the smallest

---

## What if we want to do both?

* Elastic net!

--

$$RSS + \lambda_1\sum_{j=1}^p\beta^2_j+\lambda_2\sum_{j=1}^p|\beta_j|$$

--

.question[
What is the $\ell_1$ part of the penalty?
]

--

.question[
What is the $\ell_2$ part of the penalty
]

---

## Elastic net

$$RSS + \lambda_1\sum_{j=1}^p\beta^2_j+\lambda_2\sum_{j=1}^p|\beta_j|$$

.question[
When will this be equivalent to Ridge Regression?
]

---

## Elastic net

$$RSS + \lambda_1\sum_{j=1}^p\beta^2_j+\lambda_2\sum_{j=1}^p|\beta_j|$$

.question[
When will this be equivalent to Lasso?
]

---

## Elastic Net

$$RSS + \lambda_1\sum_{j=1}^p\beta^2_j+\lambda_2\sum_{j=1}^p|\beta_j|$$


* The $\ell_1$ part of the penalty will generate a **sparse** model (shrink some $\beta$ coefficients to exactly 0)
--

* The $\ell_2$ part of the penalty removes the limitation on the number of variables selected (can be $>n$ now)
--

.question[
How do you think $\lambda_1$ and $\lambda_2$ are chosen?
]