---
title: "Decision trees - Regression tree example"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan <i>adapted from slides by Hastie & Tibshirani</i>
</span>
</div> 

```{r, include = FALSE}
library(tidyverse)
library(ISLR)
library(tidymodels)
set.seed(1)
```

---

class: center, middle

## The baseball example

---

## 1. Randomly divide the data in half, 132 training observations, 131 testing

```{r, echo = FALSE}
baseball <- Hitters[!is.na(Hitters$Salary), ]
```

```{r}
set.seed(77)
baseball_split <- initial_split(baseball, prop = 0.5)
baseball_train <- training(baseball_split)
```

---

## 2. Create cross-validation object for 6-fold cross validation

```{r}
baseball_cv <- vfold_cv(baseball_train, v = 6)
```

---

## 3. Create a model specification that tunes based on complexity, $\alpha$

```{r}
tree_spec <- decision_tree(
  cost_complexity = tune(), 
  tree_depth = 10,
  mode = "regression") %>%
  set_engine("rpart")
```

--

.question[
What is my tree depth for my "large" tree?
]

---

## 4. Fit the model on the cross validation set

.small[ 
```{r, cache = TRUE}
grid <- expand_grid(cost_complexity = seq(0.01, 0.05, by = 0.01))
model <- tune_grid(tree_spec,
                   Salary ~ Hits + Years + PutOuts + RBI + Walks + Runs,
                   grid = grid,
                   resamples = baseball_cv)
```
]

--

.question[
What $\alpha$s am I trying?
]

---

## 5. Choose $\alpha$ that minimizes the RMSE

```{r}
model %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  arrange(mean)
```

---

## 5. Choose $\alpha$ that minimizes the RMSE

```{r, eval = FALSE}
model %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  arrange(mean)
```

```{r}
model %>%
  select_best(metric = "rmse") #<<
```

---

## 5. Choose $\alpha$ that minimizes the RMSE

```{r, eval = FALSE}
model %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  arrange(mean)
```

```{r}
final_complexity <- model %>%
  select_best(metric = "rmse") %>%
  pull() #<<
```

---

## 6. Fit the final model

.small[
```{r}
final_spec <- decision_tree(
  cost_complexity = final_complexity,
  tree_depth = 10,
  mode = "regression") %>%
  set_engine("rpart")

final_model <- fit(final_spec,
                   Salary ~ Hits + Years + PutOuts + RBI + Walks + Runs,
                   data = baseball_train)
```
]

---

## Final tree

```{r, echo = FALSE}
rpart.plot::rpart.plot(final_model$fit, roundint = FALSE)
```

--

.question[
How many terminal nodes does this tree have?
]

---

## Calculate RMSE on the test data

```{r}
baseball_test <- testing(baseball_split)
final_model %>%
  predict(new_data = baseball_test) %>%
  bind_cols(baseball_test) %>%
  metrics(truth = Salary, estimate = .pred)
```

---

class: inverse

`r countdown::countdown(minutes = 10)`

## <i class="fas fa-laptop"></i> `AE 05 - Regression trees`

1. Find starter files, beginning with `appex-05-regression-trees` on GitHub and pull into RStudio
2. Complete the exercises
3. Knit, Commit, Push frequently. Be sure to have the final results pushed to GitHub by **April 3 at noon**
