---
title: "Plotting Decision trees"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

class: center, middle

`r countdown::countdown(minutes = 5)`

## `r emo::ji("book")` Take 06 Assessment in Canvas

---

## Plotting decision trees

There are several R packages that assist with tree plotting

* `rpart.plot`
* `partykit`
* `rattle`

---

## Where to find out more about packages

1. Vignettes
2. Journal Article (R Journal is great)
3. Rstudio Community
4. StackOverflow
5. Twitter

---


class: inverse

`r countdown::countdown(minutes = 5)`

## <i class="fas fa-laptop"></i> `Learn about partykit`

Google `partykit` to find out how more about it (hint: use key words like `rstats` and `decision tree plots`)

---

## `rpart.plot`

We're going to focus on `rpart.plot`, but feel free to try the others!

--

```{r, eval = FALSE}
install.packages("rpart.plot")
```

--

```{r, message = FALSE, warning = FALSE}
library(rpart.plot)
```

---

## `rpart.plot`

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(ISLR)
library(tidymodels)
baseball <- Hitters[!is.na(Hitters$Salary), ]
```

.small[
```{r}
tree_spec <- decision_tree(
  cost_complexity = 0.1,
  tree_depth = 10,
  mode = "regression") %>%
  set_engine("rpart")

model <- fit(tree_spec,
                   Salary ~ Hits + Years + PutOuts + RBI + Walks + Runs,
                   data = baseball)
```
]

--

```{r, eval = FALSE}
rpart.plot(model$fit,
           roundint = FALSE)
```

---

## `rpart.plot`


.small[
```{r}
tree_spec <- decision_tree(
  cost_complexity = 0.1,
  tree_depth = 10,
  mode = "regression") %>%
  set_engine("rpart")

model <- fit(tree_spec,
                   Salary ~ Hits + Years + PutOuts + RBI + Walks + Runs,
                   data = baseball)
```
]

```{r, eval = FALSE}
rpart.plot(model$fit, #<<
           roundint = FALSE)
```

---

## `rpart.plot`


.small[
```{r}
tree_spec <- decision_tree(
  cost_complexity = 0.1,
  tree_depth = 10,
  mode = "regression") %>%
  set_engine("rpart")

model <- fit(tree_spec,
                   Salary ~ Hits + Years + PutOuts + RBI + Walks + Runs,
                   data = baseball)
```
]

```{r, eval = FALSE}
rpart.plot(model$fit, 
           roundint = FALSE) #<<
```

---

## `rpart.plot`

```{r}
rpart.plot(model$fit, 
           roundint = FALSE) #<<
```

---


class: inverse

`r countdown::countdown(minutes = 10)`

## <i class="fas fa-laptop"></i> `AE 05 - Regression trees`

1. Open your application exercise (05) from last week
2. Install `rpart.plot`
3. Create a plot from your final decision tree
3. Knit, Commit, Push. Be sure to have the final results pushed to GitHub by **April 8 at noon**

