---
title: "Random Forests"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan <i>adapted from slides by Hastie & Tibshirani</i>
</span>
</div> 

```{r, include = FALSE}
library(tidyverse)
library(ISLR)
library(tidymodels)
set.seed(1)
```

---


class: center, middle

## The heart disease example

_Recall that we are predicting whether a patient has heart disease from 13 predictors_
---

## 1. Randomly divide the data in half, 149 training observations, 148 testing

```{r, echo = FALSE, message = FALSE, warning = FALSE}
heart <- read_csv("data/heart.csv")
heart$HD <- as.factor(heart$AHD)
heart <- heart[complete.cases(heart), ]
```

```{r}
set.seed(77)
heart_split <- initial_split(heart, prop = 0.5)
heart_train <- training(heart_split)
```

---

## 2. Create model specification

```{r, eval = FALSE}
model_spec <- rand_forest(
  mode = "classification",
  mtry = ---
) %>% 
  set_engine("ranger")
```

.question[
`mtry` here is `m`. If we are doing _bagging_ what do you think we set this to? 
]

---

## 2. Create bagging specification

```{r, eval = FALSE}
bagging_spec <- rand_forest(
  mode = "classification",
  mtry = 13 #<<
) %>% 
  set_engine("ranger")
```

--

.question[
What would we change `mtry` to if we are doing a random forest?
]

---

## 2. Create Random Forest specification

```{r}
rf_spec <- rand_forest(
  mode = "classification",
  mtry = 3 #<<
) %>% 
  set_engine("ranger")
```

--

* The default for `rand_forest` is `floor(sqrt(# predictors))` (so 3 in this case)

---

## 3. Fit the model

```{r}
model <- fit(rf_spec,
             HD ~ Age + Sex + ChestPain + RestBP + Chol + Fbs + 
               RestECG + MaxHR + ExAng + Oldpeak + Slope + Ca + Thal,               
             data = heart_train)
```

---

## 4. Examine how it looks in the test data

```{r}
heart_test <- testing(heart_split)
model %>%
  predict(new_data = heart_test) %>%
  bind_cols(heart_test) %>%
  conf_mat(truth = HD, estimate = .pred_class) %>%
  autoplot(type = "heatmap")
```


---

class: inverse

`r countdown::countdown(minutes = 10)`

## <i class="fas fa-laptop"></i> `AE 05 - Random Forests`

1. Open your application exercise (05) from last class
2. Refit your model as a _random forest_
3. Knit, Commit, Push. Be sure to have the final results pushed to GitHub by **April 10 at noon**


