---
title: "Boosted Trees in R"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

## Boosted trees in R

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidymodels)
library(tidyverse)
heart <- read_csv("data/heart.csv")
heart$HD <- as.factor(heart$AHD)
heart <- heart[complete.cases(heart), ]
```

```{r, eval = FALSE}
library(tidymodels)
```

```{r}
boost_spec <- boost_tree(  #<<
  mode = "classification", 
  tree_depth = 1, 
  trees = 1000, 
  learn_rate = 0.001, 
) %>% 
  set_engine("xgboost")  #<<
```

---

## Boosted trees in R


```{r, eval = FALSE}
library(tidymodels)
```

```{r}
boost_spec <- boost_tree( 
  mode = "classification", #<<
  tree_depth = 1, 
  trees = 1000, 
  learn_rate = 0.001, 
) %>% 
  set_engine("xgboost") 
```

* Set the `mode` as you would with a bagged tree or random forest

---

## Boosted trees in R


```{r, eval = FALSE}
library(tidymodels)
```

```{r}
boost_spec <- boost_tree( 
  mode = "classification", 
  tree_depth = 1, #<<
  trees = 1000,
  learn_rate = 0.001,
) %>% 
  set_engine("xgboost") 
```

* `tree_depth` here is the depth of each tree, let's set that to 1  

---


## Boosted trees in R


```{r, eval = FALSE}
library(tidymodels)
```

```{r}
boost_spec <- boost_tree( 
  mode = "classification", 
  tree_depth = 1, 
  trees = 1000, #<<
  learn_rate = 0.001,
) %>% 
  set_engine("xgboost") 
```

* `trees` is the number of trees that are fit, this is equivalent to `B`

---


## Boosted trees in R


```{r, eval = FALSE}
library(tidymodels)
```

```{r}
boost_spec <- boost_tree( 
  mode = "classification", 
  tree_depth = 1, 
  trees = 1000, 
  learn_rate = 0.001, #<<
) %>% 
  set_engine("xgboost") 
```

* `learn_rate` is $\lambda$

---

## Fit the model

```{r}
model <- fit(boost_spec,
             HD ~ Age + Sex + ChestPain + RestBP + Chol + Fbs + 
               RestECG + MaxHR + ExAng + Oldpeak + Slope + Ca + Thal,               
             data = heart)
```


---

class: inverse

`r countdown::countdown(minutes = 3)`

## <i class="fas fa-laptop"></i> `Boosting`

How would this code change if I wanted to tune `B` the number of bootstrapped training samples?

```{r}
boost_spec <- boost_tree( 
  mode = "classification", 
  tree_depth = 1, 
  trees = 1000, 
  learn_rate = 0.001, 
) %>% 
  set_engine("xgboost") 
```

