---
title: "Variable Importance"
author: "Dr. D'Agostino McGowan"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r child = "setup.Rmd"}
```

layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan <i>adapted from slides by Hastie & Tibshirani</i>
</span>
</div> 

---

## Variable importance

* For bagged or random forest _regression trees_, we can record the _total RSS_ that is decreased due to splits of a given predictor $X_i$ averaged over all $B$ trees

--
* A large value would indicate that that variable is _important_

---

## Variable importance

* For bagged or random forest _classification trees_ we can add up the total amount that the Gini Index is decreased by splits of a given predictor, $X_i$, averaged over $B$ trees

---

## Variable importance in R

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidymodels)
library(tidyverse)
heart <- read_csv("data/heart.csv")
heart$HD <- as.factor(heart$AHD)
heart <- heart[complete.cases(heart), ]
```

.small[
```{r}
rf_spec <- rand_forest(
  mode = "classification",
  mtry = 3
) %>% 
  set_engine(
    "ranger",
    importance = "impurity") #<<

model <- fit(rf_spec,
             HD ~ Age + Sex + ChestPain + RestBP + Chol + Fbs + 
               RestECG + MaxHR + ExAng + Oldpeak + Slope + Ca + Thal,               
             data = heart)
```
]

```{r}
ranger::importance(model$fit)
```

---

## Variable importance

.small[
```{r}
library(ranger)
importance(model$fit)
```
]

--

.small[
```{r}
var_imp <- ranger::importance(model$fit)
```
]

---

## Plotting variable importance

.small[
```{r, fig.height = 1.5}
var_imp_df <- data.frame(
  variable = names(var_imp),
  importance = var_imp
  )

var_imp_df %>%
ggplot(aes(x = variable, y = importance)) +
  geom_col()
```
]

--

.question[
How could we make this plot better?
]

---

## Plotting variable importance

.small[
```{r}
var_imp_df %>%
ggplot(aes(x = variable, y = importance)) +
  geom_col() + 
  coord_flip()
```
]

.question[
How could we make this plot better?
]

---


## Plotting variable importance

.small[
```{r}
var_imp_df %>%
  mutate(variable = factor(variable, 
   levels = variable[order(var_imp_df$importance)])) %>%
ggplot(aes(x = variable, y = importance)) +
  geom_col() + 
  coord_flip()
```
]

---


