---
title: "Lab 04 - Ridge Lasso Elastic Net"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(ISLR)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```

**Due:** Tuesday 2020-03-24 at 11:59pm

# Introduction

In this lab we are going to practice cross validation. A few reminders:

* Remember to label your chunks
* Write out your answers in full sentences, do not just rely on the R output
* Knit, commit, and push regularly (at least after every exercise)

# Getting started

- Go to our class's GitHub organization [sta-363-s20](https://github.com/sta-363-s20)
- Find the GitHub repository (which we'll refer to as "repo" going forward) for this lab, `lab-04-ridge-lasso-elastic-YOUR-GITHUB-HANDLE`. This repo contains a template you can build on to complete your assignment.
- On GitHub, click on the green **Clone or download** button, select **Use HTTPS**. Click on the clipboard icon to copy the repo URL.
- If using RStudio Cloud, go to RStudio Cloud and into the course workspace. Create a **New Project from Git Repo**. You will need to click on the down arrow next to the **New Project** button to see this option.
- If using RStudio Pro, create a new project by clicking **File > New Project** Then click **Version Control** and **Git/Github**.
- Copy and paste the URL of your assignment repo into the dialog box.
- Hit OK, and you're good to go!

# Packages

In this lab we will work with two packages: `tidyverse` which is a collection of packages for doing data analysis in a "tidy" way and `tidymodels` for statistial modeling.

Now that the necessary packages are installed, you should be able to Knit your document and see the results.

If you'd like to run your code in the Console as well you'll also need to load the packages there. To do so, run the following in the console. 

```{r eval = FALSE}
library(tidyverse) 
library(tidymodels)
```

Note that the packages are also loaded with the same commands in your R Markdown document.

# Housekeeping

## Git configuration

```{marginfigure}
Your email address is the address tied to your GitHub account and your name should be first and last name.
```

- Go to the *Terminal* pane
- Type the following two lines of code, replacing the information in the quotation marks with your info:

```{bash eval=FALSE}
git config --global user.email "your email"
git config --global user.name "your name"
```

To confirm that the changes have been implemented, run the following

```{bash eval=FALSE}
git config --global user.email
git config --global user.name
```

## Password caching

If you would like your git password cached for a week for this project, type the following in the Terminal:

```{bash eval=FALSE}
git config --global credential.helper 'cache --timeout 604800'
```

## Project name: 

Currently your project is called *Untitled Project*. Update the name of your project to be "Lab 04 - Ridge, Lasso, and Elastic Net".

# Warm up

Before we introduce the data, let's warm up with some simple exercises.

## YAML: 

Open the R Markdown (Rmd) file in your project, change the author name to your name, and knit the document.

## Commiting and pushing changes:

- Go to the **Git** pane in your RStudio. 
- View the **Diff** and confirm that you are happy with the changes.
- Add a commit message like "Update team name" in the **Commit message** box and hit **Commit**.
- Click on **Push**. This will prompt a dialogue box where you first need to enter your user name, and then your password.

# Data

For this lab, we are using a data frame currently in `music.csv`. This data frame includes 72 predictors that are components of audio files and one outcome, `lat`, the latitude of where the music originated. We are trying to predict the location of the music's origin using audio components of the music.

# Exercises

1. Set a seed of `7`. Split the `music` data into a training and test set with 50% of the data in the training and 50% in the testing. Call your training set `music_train` and your testing set `music_test`. Describe these data sets (how many observations, how many variables).


```{r split}
music <- read_csv("music.csv")
```


2. We are interested in predicting the latitude (`lat`) of the music's origin from all other variables. Fit a linear model using least squares on the training set. Report the test root mean squared error obtained.

3.  Fit a ridge regression model on the training set with $\lambda$ chosen using 10-fold cross valiation. Report the $\lambda$ chosen and explain why. Report the test root mean squared error obtained using testing portion of the initially split data frame.


4. Fit a lasso model on the training set with $\lambda$ chosen using cross valiation. Report the $\lambda$ chosen and explain why. Report the test root mean squared error obtained using testing portion of the initially split data frame.


5. Fit an elastic net model on the training set with $\lambda$ and $\alpha$ chosen using cross valiation. Report the $\lambda$ chosen and explain why. Report the test root mean squared error obtained using testing portion of the initially split data frame.


6. Comment on the results obtained. How accurately can we predict the latitude of where the music originated? Is there much difference among the test errors resulting from these four approaches?
